---
title: "Homework #5: Density Estimation" 
author: "**Jie Tang**"
date: "Due: Tue Mar 22 | 10:30 am"
output: R6018::homework
editor_options:
  chunk_output_type: console
---

**SYS 4582/6018 | Spring 2022 | University of Virginia **

*******************************************
```{r config, echo=FALSE}
source(system.file("config/hw_config.R", package="R6018")) # knitr settings
options(dplyr.summarise.inform = FALSE)  # ignore dplyr message about grouping
```


# Required R packages and Directories

::: {.solution}
```{r packages, message=FALSE, warning=FALSE}
data.dir = 'https://mdporter.github.io/SYS6018/data/' # data directory
batdata.dir = ' https://mdporter.github.io/SYS6018/data//geo_profile.csv'
library(R6018)     # functions for SYS 6018
library(ks)        # functions for KDE
library(tidyverse) # functions for data manipulation   
```
:::



# Problem 1 Geographic Profiling

Geographic profiling, a method developed in criminology, can be used to estimate the [home location (roost) of animals](https://www.sciencedirect.com/science/article/pii/S0022519305004157) based on a collection of sightings. The approach requires an estimate of the distribution the animal will travel from their roost to forage for food. 

A sample of $283$ distances that pipistrelle bats traveled (in meters) from their roost can be found at: 

- **Bat Data**: <`r file.path(data.dir, 'geo_profile.csv')`>

One probability model for the distance these bats will travel is:
\begin{align*}
f(x; \theta) = \frac{x}{\theta} \exp \left( - \frac{x^2}{2 \theta} \right)
\end{align*}
where the parameter $\theta > 0$ controls how far they are willing to travel. 


## a. Derive the MLE for $\theta$ (i.e., show the math). 

::: {.solution}
\begin{align*}
f(x; \theta) = \frac{x}{\theta} \exp \left( - \frac{x^2}{2 \theta} \right)
f(x; \theta) = 
\end{align*}
:::


## b. What is the MLE of $\theta$ for the bat data? (Use results from a, or use computational methods.) 

::: {.solution}
```{r}
batdata = read.csv('geo_profile.csv', header=FALSE)
head(batdata)
x = batdata$V1
#mle
mle = sum(x^2)/(2*length(x)) 
```

:::

## c. Using the MLE value of $\theta$ from part b, compute the estimated density at a set of evaluation points between 0 and 8 meters. Plot the estimated density.

::: {.solution}
```{r}
theta = mle
x = seq(0, 8, length=200)
outcome = tibble(x = x, y = (x/theta) * exp(-x^2/(2*theta)))
ggplot(outcome, aes(x, y)) + geom_line()

#ggplot(outcome, aes(x, y)) + geom_density(aes(x, y), bw=1, color="red")

```

:::

## d. Estimate the density using KDE. Report the bandwidth you chose and produce a plot of the estimated density. 

::: {.solution}
```{r}
tibble(x) %>% 
  ggplot(aes(x)) + 
  geom_histogram(aes(y=stat(density)), alpha=.8) + # add histogram
  geom_density(bw=0.35, color="red")



```

:::




## e. Which model do you prefer, the parametric or KDE? 

::: {.solution}
Add solution here
:::




# Problem 2: Interstate Crash Density

Interstate 64 (I-64) is a major east-west road that passes just south of Charlottesville. Where and when are the most dangerous places/times to be on I-64? The crash data (link below) gives the mile marker and fractional time-of-week for crashes that occurred on I-64 between mile marker 87 and 136 in 2016. The time-of-week data takes a numeric value of *\<dow\>.\<hour/24\>*, where the dow starts at 0 for Sunday (6 for Sat) and the decimal gives the time of day information. Thus `time=0.0417` corresponds to Sun at 1am and `time=6.5` corresponds to Sat at noon. 

- **Crash Data**: <`r file.path(data.dir, 'crashes16.csv')`>


## a. Extract the crashes and make a scatter plot with mile marker on x-axis and time on y-axis. 

::: {.solution}
```{r}
crashdata = read.csv("crashes16.csv")
head(crashdata)
ggplot(crashdata, aes(mile, time)) + geom_point()
```
:::



## b. Use KDE to estimate the *mile marker* density. 

- Report the bandwidth. 
- Plot the density estimate.

::: {.solution}
```{r}
#The bandwidth I chose is 2.35,  by the range of value divided by 20
library(Hmisc)
describe(crashdata$mile)
bw = (134 - 87) / 20
ggplot(crashdata, aes(mile)) + 
  geom_density(bw=bw)
```

:::

## c. Use KDE to estimate the temporal *time-of-week* density. 

- Report the bandwidth. 
- Plot the density estimate. 

::: {.solution}
```{r}
#The bandwidth I chose is 0.4325, by the range of value divided by 20
describe(crashdata$time)
bw = (6.9 - 0.05) / 20
ggplot(crashdata, aes(time)) + 
  geom_density(bw=bw)
```

:::

## d. Use KDE to estimate the bivariate mile-time density. 

- Report the bandwidth parameters.
- Plot the bivariate density estimate. 

::: {.solution}
```{r}

```

:::


## e. Based on the estimated density, approximate the most dangerous place and time to drive on this strech of road. Identify the mile marker and time-of-week pair. 

::: {.solution}
Add solution here
:::


		